### 2024-07-01 causal-inference.qmd
20240701 - ggdag

```{r}
library("dagitty")
```

@@

plot(dagitty::graphLayout(XCausesY))
```

@@

plot(dagitty::graphLayout(YCausesX))
```

@@

>>>>

```{r}
library("dagitty")
library("ggdag")
```

@@

plot(dagitty::graphLayout(XCausesY))

dagitty::impliedConditionalIndependencies(XCausesY)
```

```{r}
#| eval: false

XCausesY_alt <- ggdag::dagify(
  Y ~ X
)

plot(XCausesY_alt)
ggdag(XCausesY_alt)
```

@@

plot(dagitty::graphLayout(YCausesX))

dagitty::impliedConditionalIndependencies(YCausesX)
```

```{r}
#| eval: false

YCausesX_alt <- ggdag::dagify(
  X ~ Y
)

plot(YCausesX_alt)
ggdag(YCausesX_alt)
```

@@

```{r}
#| eval: false

ZCausesXandY_alt <- ggdag::dagify(
  X ~ Z,
  Y ~ Z,
  X ~~ Y
)

plot(ZCausesXandY_alt)
ggdag(ZCausesXandY_alt)
```

Consider another example in @fig-dag:

```{r}
#| label: fig-dag
#| fig-cap: "Causal Diagram (Directed Acyclic Graph)."

mediationDag <- dagitty::dagitty("dag{
  X -> M
  X -> A
  M -> Y
  Y -> Z
  A <-> M
}")

plot(dagitty::graphLayout(mediationDag))

dagitty::impliedConditionalIndependencies(mediationDag)

dagitty::adjustmentSets(
  mediationDag, 
  exposure = "X",
  outcome = "Z",
  effect = "total")
```

```{r}
#| eval: false

mediationDag_alt <- ggdag::dagify(
  M ~ X,
  Y ~ M,
  Z ~ Y,
  A ~ X,
  A ~~ M
)

plot(mediationDag_alt)
ggdag(mediationDag_alt)
```
